Heritability

### 2021-06-29 Manuscript/manuscript.Rmd
update manuscript with new analyses and update makefile

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(captioner)
library(broom)
library(broman)
library(morris)
library(metafor)
figures <- captioner()
figures("deviance", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment. Deviance is calculated as the mean of P within each passage minus the mean of N within each passage. Errorbars are the standard error of the difference between P and N for each passage. The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations. The gray ribbon is the 95% confidence interval for the regression.")
figures("herit", "Regression of offspring CH~4~ flux on mid-parent CH~4~ flux. Mid-parent is the mean of jars selected to inoculate the next passage. The mid-offsprings are the mean of all twelve jars produced in one passage. The best fit line is an ordinary least squares regression line and the gray ribbon is the 95% confidence interval for the regression coefficient.")

@@

```{r dev_model}
fluxes <- read_tsv('../Output/fluxes.tsv')
deviance <- read_tsv('../Output/deviance.tsv')
rma_output <- readRDS('../Output/response.rds')
dev_fit <- readRDS('../Output/dev_fit.rds')
dev_pred <- predict(dev_fit)

flux_per_psg <-
  fluxes %>% 
  group_by(passage) %>% 
  summarize(mean = mean(estimate), se = morris::se(estimate))

passages <- length(deviance$passage)

Methane oxidation rate increased as a response to selection in the Positive selection treatment (`r figures("deviance", display="cite")`). 
This can be seen in the increase in deviation between the Positive and Neutral selection lines over the course of the experiment (z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`). 
Mean CH~4~ oxidation rate for the Positive treatment increased by `r myround(dev_fit$beta[2], 3)` $\pm$ `r myround(dev_fit$se[2], 3)` per passage relative to the Neutral treatment. The mean flux across all jars in all treatments in the first passage was `r myround(flux_per_psg[1, 2], 3)` $\pm$ `r myround(flux_per_psg[1, 3], 3)` so this represents a doubling of CH~4~ consumption in every passage due to selection.

```{r dev_fig, fig.width=4, fig.height=3, fig.cap=figures("deviance")}
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
  geom_pointrange(color = 'darkorange2') + 
      geom_segment(aes(x = 1, xend = passages, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*passages), color = 'darkorange2', size = 1) +
  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
  theme_bw()

@@

```{r herit_model}
heritability <- read_tsv('../Output/heritability.tsv')

her_fit <- lm(offspring ~ parental, data = heritability)
her_sum <- tidy(her_fit)
her_aov <- tidy(anova(her_fit))

```{r inbreeding_coef}
N = 12 # population size
T = 5  # number of passages (generations)
F = 1 - (1 - 1 / (2 * N)) ^ T
```

Methane oxidation rate was heritable between consecutive passages (`r figures("herit", display="cite")`). Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)` (t = `r myround(her_sum[2, 4], 2)`, p = `r myround(her_sum[2, 5], 3)`). With an inbreeding coefficient of `r myround(F, 2)`,

```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
ggplot(heritability, aes(x = parental, y = offspring)) +
  geom_point(aes(color = treat)) +
  stat_smooth(method = 'lm', formula = y ~ x, color = "black") + 
  labs(x = "Parental Mean Flux (-k)", y = "Offspring Mean Flux (-k)") +
    scale_color_manual(name = "Selection\nTreatment", labels = c('Neutral', 'Positive'), values = c("gray60", "darkorange2")) +
  theme_bw()

```{r}
R = dev_fit$beta[2]
h2 = pull(her_sum[2, 2])
S = R/h2
```

>>>>

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r initialize_packages}
library(tidyverse)
library(captioner)
library(broom)
library(broman)
library(morris)
#library(metafor)
```
```{r load_data}
heritability <- read_tsv('../Output/heritability.tsv')
fluxes <- read_tsv('../Output/fluxes.tsv')

load(file = '../Output/response_model.Rdata')
load(file = '../Output/herit_model.Rdata')

source('../R/functions.R')

```{r figure_captions, include=FALSE}

figures <- captioner()
figures("response", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment. Deviance is calculated as the mean of P within each passage minus the mean of N within each passage. Errorbars are the standard error of the difference between P and N for each passage. The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations. The gray ribbon is the 95% confidence interval for the regression.")
figures("herit", "Regression of offspring CH~4~ flux on mid-parent CH~4~ flux. Mid-parent is the mean of jars selected to inoculate the next passage. The mid-offsprings are the mean of all twelve jars produced in one passage. The best fit line is an ordinary least squares regression line and the gray ribbon is the 95% confidence interval for the regression coefficient.")

@@

The intercept for the Neutral treatment was significantly different from zero
at `r myround(tidy(response_model)[[1, 2]], 2)` (SE = `r
myround(tidy(response_model)[[1, 3]], 2)`, t = `r myround(tidy(response_model)[[1,
4]], 2)`, p = `r myround(tidy(response_model)[[1, 5]], 2)`). The Positive
treatment started at a significantly lower methane oxidation rate than the
Neutral treatment with a y-intercept of `r myround(tidy(response_model)[[3, 2]],
2)` relative to Neutral (SE = `r myround(tidy(response_model)[[3, 3]], 2)`, t = `r
myround(tidy(response_model)[[3, 4]], 2)`, p = `r myround(tidy(response_model)[[3,
5]], 2)`). There was no change in methane oxidation rate of the Neutral
treatment over the five passages (slope = `r myround(tidy(response_model)[[2, 2]],
2)`, SE = `r myround(tidy(response_model)[[2, 3]], 2)`, t = `r
myround(tidy(response_model)[[2, 4]], 2)`, p = `r myround(tidy(response_model)[[2,
5]], 2)`). By contrast, there was a significant increase in the Positive
treatment of `r myround(tidy(response_model)[[4, 2]], 2)` per passage which is
approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane
oxidation rate per passage (SE = `r myround(tidy(response_model)[[4, 3]], 2)`, t =
`r myround(tidy(response_model)[[4, 4]], 2)`, p = `r myround(tidy(response_model)[[4,
5]], 2)`).

Methane oxidation rate increased as a response to selection in the Positive
selection treatment (`r figures("deviance", display="cite")`).  This can be
seen in the increase in deviation between the Positive and Neutral selection
lines over the course of the experiment (z = , p = ).  Mean CH~4~ oxidation
rate for the Positive treatment increased by  $\pm$ per passage relative to the
Neutral treatment. The mean flux across all jars in all treatments in the first
passage was  $\pm$  so this represents a doubling of CH~4~ consumption in every passage due
to selection.

```{r response_figure, fig.width=4, fig.height=3, fig.cap=figures("response")}
ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)

plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)

@@

Here, I fit two models. The first is simply offspring flux on selected parent
flux, ignoring treatment. There is no correlation. Next, I added treatment and the interaction between treatment and parental
flux. Here, there is a significant correlation between parent and offspring for
both the Neutral treatment `r print_model(herit_model, 2)` and the Positive
treatment `r print_model(herit_model, 4)`. Notably, the sign of the effect is
reversed: negative for the Neutral treatment and positive for the Positive
treatment.

The negative heritability of methane oxidation rate in the Neutral selection
treatment may result from selection on alternative characters in those
populations. The strong, positive heritability in the Positive treatment ... .
It seems strange that the Methane oxidation rate was heritable between
consecutive passages (`r figures("herit", display="cite")`). Specifically, the
narrow-sense heritability, calculated as the slope of the mid-parent and
mid-offspring, was $h^2$ =  $\pm$  (t = , p = ). With an inbreeding coefficient of ,

```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
plot_herit(heritability)
```

### 2021-06-29 R/03-response.Rmd
update manuscript with new analyses and update makefile

This script intakes the flux data from script 01. It then fits a linear model to the flux data over time per treatment and tests a difference of slope between the positive and the neutral treatment. Finally, it 
```{r initialize_packages, message=FALSE}
library(tidyverse)
library(morris)
library(broom)
library(knitr)
#library(broman)
```

@@

fluxes <-
  fluxes %>% 
  mutate(passage = passage - 1)
```

@@

```{r response_model}
# Fit full interaction model
inter_model <- lm(log_ch4 ~ passage * treat, data = fluxes)

@@

# Back calculate the slope of the positive treatment
pos_slope <- 10^inter_model$coef["passage:treatp"]

@@

Here I fit two models with flux as the dependent variable and passage and treatment as the independent variables. The first is the interaction model, which includes an interaction between passage and treatment. The second is an additive model that lacks this interaction. Here is the model comparison using `anova`, which indicates the full model is somewhat better:

```{r compare_models}
# Compare interaction to no interaction
compare_mods <- anova(inter_model, no_treat_model)
kable(tidy(compare_mods))
```

@@

```{r inter_model}
kable(tidy(inter_model))

The intercept for the Neutral treatment was significantly different from zero at `r round(tidy(inter_model)[[1, 2]], 2)` (SE = `r round(tidy(inter_model)[[1, 3]], 2)`, t = `r round(tidy(inter_model)[[1, 4]], 2)`, p = `r round(tidy(inter_model)[[1, 5]], 2)`). The Positive treatment started at a significantly lower methane oxidation rate than the Neutral treatment with a y-intercept of `r round(tidy(inter_model)[[3, 2]], 2)` relative to Neutral (SE = `r round(tidy(inter_model)[[3, 3]], 2)`, t = `r round(tidy(inter_model)[[3, 4]], 2)`, p = `r round(tidy(inter_model)[[3, 5]], 2)`). There was no change in methane oxidation rate of the Neutral treatment over the five passages (slope = `r round(tidy(inter_model)[[2, 2]], 2)`, SE = `r round(tidy(inter_model)[[2, 3]], 2)`, t = `r round(tidy(inter_model)[[2, 4]], 2)`, p = `r round(tidy(inter_model)[[2, 5]], 2)`). By contrast, there was a significant increase in the Positive treatment of `r round(tidy(inter_model)[[4, 2]], 2)` per passage which is approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane oxidation rate per passage (SE = `r round(tidy(inter_model)[[4, 3]], 2)`, t = `r round(tidy(inter_model)[[4, 4]], 2)`, p = `r round(tidy(inter_model)[[4, 5]], 2)`).

>>>>

This script takes in the flux data from script 01. It then plots the raw data
and the log-transformed data. Finally, it fits a linear model to the flux data
over time per treatment and tests a difference of slope between the positive
and the neutral treatment.

```{r initialize_packages, message=FALSE}
library(tidyverse)
library(morris)
library(broom)
library(knitr)
library(broman)
```

@@

# Shift the passage numbers so the first passage becomes "Passage 0"
fluxes <-
  fluxes %>% 
  mutate(passage = passage - 1)
```

@@

```{r response_model}
# Fit full interaction model
response_model <- lm(log_ch4 ~ passage * treat, data = fluxes)

@@

# Back calculate the slope of the positive treatment
pos_slope <- 10^response_model$coef["passage:treatp"]

@@

Here I fit two models. The first is the full model which includes passage number, treatment, and the interaction between the two. The second model includes only passage with no effect of treatment. Comparing the models using `anova`

```{r compare_models}
# Compare interaction to no interaction
compare_mods <- anova(response_model, no_treat_model)
kable(tidy(compare_mods))
```

@@

```{r print_model}
kable(tidy(response_model))
```

```{r save_model}
save(response_model, pos_slope, file = '../Output/response_model.Rdata')
```

The intercept for the Neutral treatment was significantly different from zero
at `r myround(tidy(response_model)[[1, 2]], 2)` (SE = `r
myround(tidy(response_model)[[1, 3]], 2)`, t = `r myround(tidy(response_model)[[1,
4]], 2)`, p = `r myround(tidy(response_model)[[1, 5]], 2)`). The Positive
treatment started at a significantly lower methane oxidation rate than the
Neutral treatment with a y-intercept of `r myround(tidy(response_model)[[3, 2]],
2)` relative to Neutral (SE = `r myround(tidy(response_model)[[3, 3]], 2)`, t = `r
myround(tidy(response_model)[[3, 4]], 2)`, p = `r myround(tidy(response_model)[[3,
5]], 2)`). There was no change in methane oxidation rate of the Neutral
treatment over the five passages (slope = `r myround(tidy(response_model)[[2, 2]],
2)`, SE = `r myround(tidy(response_model)[[2, 3]], 2)`, t = `r
myround(tidy(response_model)[[2, 4]], 2)`, p = `r myround(tidy(response_model)[[2,
5]], 2)`). By contrast, there was a significant increase in the Positive
treatment of `r myround(tidy(response_model)[[4, 2]], 2)` per passage which is
approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane
oxidation rate per passage (SE = `r myround(tidy(response_model)[[4, 3]], 2)`, t =
`r myround(tidy(response_model)[[4, 4]], 2)`, p = `r myround(tidy(response_model)[[4,
5]], 2)`).

### 2021-06-29 R/04-heritability.Rmd
update manuscript with new analyses and update makefile

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

This script imports the heritability data. This dataset includes the mean flux withn treatment*generation for 1) all parents, 2) selected parents, and 3) offspring. Then, it plots offspring on the selection parents with the two treatments combined and then separately for the two treatments.

@@

```{r test_treatment}
fit_no_treat <- lm(offspring ~ selected, data = heritability)
kable(tidy(fit_no_treat))
fit_treat <- lm(offspring ~ selected * treat, data = heritability)
kable(tidy(fit_treat))
kable(tidy(anova(fit_no_treat, fit_treat)))
```

Here, I fit two models. The first is simply offspring flux on selected parent flux, ignoring treatment. There is no correlation `r print_model(fit_no_treat, 2)`. Next, I added treatment and the interaction between treatment and parental flux. Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_model(fit_treat, 2)` and the Positive treatment `r print_model(fit_treat, 4)`. Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.

@@

>>>>

---
title: "Heritability"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

This script imports the heritability data. This dataset includes the mean flux
withn treatmentxgeneration for 1) all parents, 2) selected parents, and 3)
offspring. Then, it plots offspring on the selection parents with the two
treatments combined and then separately for the two treatments.

@@

```{r test_treatment}
fit_no_treat <- lm(offspring ~ selected, data = heritability)
kable(tidy(fit_no_treat))
herit_model <- lm(offspring ~ selected * treat, data = heritability)
kable(tidy(herit_model))
kable(tidy(anova(fit_no_treat, herit_model)))
```

Here, I fit two models. The first is simply offspring flux on selected parent flux, ignoring treatment. There is no correlation `r print_model(fit_no_treat, 2)`. Next, I added treatment and the interaction between treatment and parental flux. Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_model(herit_model, 2)` and the Positive treatment `r print_model(herit_model, 4)`. Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.

@@

```{r save_heritability}
save(herit_model, file = '../Output/herit_model.Rdata')
```
